Reinforcement Learning for Options on Target Volatility Funds

### 2021-05-26 tex/targetvol.tex
Paper
add second formulation of reward function, add thanks, terminated LV

\usepackage[]{hyperref}
\hypersetup{
	colorlinks=true,breaklinks=true,
	urlcolor= mydarkblue,linkcolor= mydarkblue,citecolor= mydarkgreen,
	pdfauthor={Daluiso, R., Nastasi, E., Pallavicini, A., Polo, S.},
}
\usepackage[backend=biber,style=authoryear-comp,maxcitenames=2,sorting=nyt,natbib=true,sortcites=false]{biblatex}
\addbibresource{Bib.bib}
%\usepackage{natbib}
%\setcitestyle{authoryear,open={(},close={)}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{remark}{Remark}[section]
\newtheorem{example}[remark]{Example}

@@

\author{
	Roberto Daluiso\thanks{Intesa Sanpaolo Milan, roberto.daluiso@intesasanpaolo.com}
	\and 
	Emanuele Nastasi\thanks{Marketz S.p.A., emanuele.nastasi@marketz.eu}
	\and Andrea Pallavicini\thanks{Imperial College London and Intesa Sanpaolo Milan, a.pallavicini@imperial.ac.uk} 
	\and
	Stefano Polo \thanks{Intesa Sanpaolo Milan, stefano.polo@intesasanpaolo.com} \thanks{We wish to thank the Italian computing centre Cineca, which approved our ISCRA C project and provided us the high-computing resources of Marconi100 for the numerical simulations of this work.}% indicates same as 2nd thanks
}

@@

\begin{abstract}
\noindent In recent years Reinforcement Learning (RL) techniques are gaining popularity in the field of quantitative finance since they are proving to be an efficient way to solve  high-dimensional optimal control problems. 
Our research project is devoted to apply RL to price derivative contracts on target volatility strategies (TVSs), portfolios of risky assets and a risk-free one dynamically rebalanced in order to keep the realized volatility of the portfolio on a certain level. The uncertainty in the TVS risky portfolio composition along with the difference in hedging costs for each component requires to solve a stochastic control problem to evaluate the option prices. The topic of hedging costs is a novelty never dealt in the TVS literature and we provide a formal description of the entire control problem. We tackle the problem by implementing a Reinforcement Learning algorithm to determine the optimal risky portfolio composition leading to the most conservative option price.  We investigate the problem for two models of the risky asset dynamics: time-dependent Black \& Scholes and Local Volatility. In the first case we prove the existence of an analytical solution of the problem; a result that we use as benchmark to perform a series of fine-tuning of the hyper-parameters of the RL algorithm. At the end we provide numerical results for the Local Volatility model, for which an \textit{a priori} solution is not available.
\end{abstract}

\smallskip
\noindent
\textbf{JEL classification codes:} C63, C45, C61, G11, G13.\\
\textbf{AMS classification codes:} 65C05, 91G20, 91G60.\\
\textbf{Keywords:} Reinforcement Learning, Hedging Costs, Funding Costs, Proximal Policy Optimization, Target Volatility, Asset Allocation. 
\newpage
\tableofcontents
\newpage
\pagestyle{myheadings} \markboth{}{{\footnotesize Daluiso, Pallavicini, Nastasi, Polo, Reinforcement Learning for Options on Target Volatility Funds}}
\section{Introduction}
In the recent years portfolio managers are exposed to very low interest rates and quickly changing market volatilities. An effective solution to control risks under such environment is given by target volatility strategies (TVSs) (also known as constant volatility targeting) which are able to preserve the portfolio at a predetermined level of volatility. A TVS is  a  portfolio  of  risky  assets  (typically  equities)  and  a  risk-free  asset dynamically re-balanced with the aim of maintaining the overall portfolio volatility level closed to some target value. This products were initially offered in the Asian markets, see for instance the reports of \textcite{Chew} and \textcite{XUE} which highlight the pros and cons for investors, to be adopted in the following years in many other markets in North America and Europe as depicted in \textcite{Morrison}.

@@

In this contribution we face a new topic never dealt with in the TVS literature: the funding costs coming from hedging the risky assets underlying the TVS. We consider the point of view of a bank selling a call option to a portfolio manager as protection on the capital invested in a TVS. The portfolio manager has the freedom of changing the relative weights of the risky assets during the life of the TVS. Since the risky assets have different hedging costs, the bank shall adjust the price of the protection to include them in the worst-case scenario, i.e. the most expensive strategy from the financing costs point of view. Hence, the pricing problem becomes a dynamical control problem over the risky portfolio composition. In our contribution we describe the dynamical control problem, and we derive an analytical solution in the simple case of the risky assets driven by a Black-Scholes (BS) model \citep{BlackScholes}. Then, we tackle the problem in the general case of a local volatility (LV) model \citep{Dupire94pricingwith,LVfirst} by using Reinforcement Learning (RL) algorithms, in particular the Proximal Policy Optimization developed in \textcite{ppo}.
\\

The paper is organized ad follows. In \Cref{sec:TVS} we describe the dynamics of a TVS in presence of valuation adjustments. Then, in \Cref{sec:Derivative} we study the case of derivative contracts on TVSs, we describe the dynamical control problem giving the expression of the Hamilton-Jacobi-Bellman equation and we derive the analytical results for plain vanilla options under BS dynamics. In \Cref{sec:RL} we illustrate how we have applied RL to solve the dynamic control problem, giving a description of the algorithm we have built. We conclude the paper with \Cref{sec:Numerical_results} where we present the numerical results obtained in this work for the Black and Scholes model and the local volatility (LV) one.

@@

where $\Phi$ is the payoff function of the derivative. We start by introducing the Markovian projection of the dynamics followed by $I_t$. We name it $I_t^{\text{MP}}$, and we get by applying the Gy\"ongy Lemma

@@

In a more general settings we are not able to find an explicit solution. A proof of the existence of the solution in a general setting is missing. This is a stochastic optimal control problem where by homogeneity we can suppose that $\alpha_t$ lives in a compact domain, namely (a subset of) the unit simplex. A least if $r_t$, $\mu^i_t$ and $\nu^i_t$ are (uniformly) bounded, and if the eigenvalues of $\nu_t$ are (uniformly) bounded away from zero, then drift and diffusion should be uniformly Lipschitz in $\alpha_t$, and classical theorems should exist. 

@@

\begin{proposition}
When the underlying securities follow a Black and Scholes (BS) model with deterministic rates, the optimal strategy for a non-decreasing European payoff consists in minimizing the local drift function, independently of the current state $I_t$
\begin{equation}
    \alpha^*(t) \coloneqq \argmin_\alpha \frac{\alpha \cdot \mu(t)}{\|\alpha \cdot \nu(t) \|}.
\label{eq:BS_optimal_strategy}
\end{equation}
Analogously, the optimal strategy for a non-increasing European payoff consists in maximizing the local drift function:
\begin{equation}
    \alpha^*(t) \coloneqq \argmax_\alpha \frac{\alpha \cdot \mu(t)}{\|\alpha \cdot \nu(t) \|}.
\label{eq:BS_optimal_strategy_put}
\end{equation}
\end{proposition}
The absence of stochastic elements in \Cref{eq:BS_optimal_strategy} makes the optimal strategy known \textit{a priori} with no simulation needed; in fact one can solve the optimization problem once for all $\forall t \in \mathcal{T}$ just looking at the market data $\mu(t)$ and $\nu(t)$ for the securities. Once $\alpha^*$ is known, then one can price the payoff by the following BS formula
\begin{equation}
	V_0^{\text{BS}} = BS(F^{\text{TVS}}(0,T;\alpha^*,K), K, T, \bar{\sigma}, D(0,T;\zeta)),
\label{eq:bs_closed_solution}\end{equation}
where $F^{\text{TVS}}(t,T;\alpha,K)$ is the TVS forward curve defined by
\begin{equation}
	F^{\text{TVS}}(t,T;\alpha, K) = I_t \exp\left[\int_t^T\left(r(u)-\ell_\alpha \left(u,K\right)\right)du\right],
\end{equation}
while $BS(F,K,T,\sigma,D)$ is the standard BS formula for an European option with forward curve $F$, strike $K$, time to maturity $T$, volatility $\sigma$ and discount factor $D$.  
\subsubsection{Black's Closed form solution}
In absence of constraints on the allocation strategy, we are able to derive a closed form solution to the BS problem \eqref{eq:BS_optimal_strategy}
\begin{lemma}
	Let be $\mu, \alpha \in \mathbb{R}^n$, $\nu \in \mathbb{R}^{n\times n}$ be a full rank matrix and $\Sigma\coloneqq \nu \tr{\nu}$. Then the closed solution of the optimization problem \eqref{eq:BS_optimal_strategy} is 
	\begin{equation}
			\alpha^* = - \frac{\Sigma^{-1} \cdot \mu}{\|(\Sigma^{-1} \cdot \mu)\cdot \nu\|}
	\end{equation}
\end{lemma}
\begin{proof}
Since the argument of the minimum \eqref{eq:BS_optimal_strategy} is zero-homogeneous, then we can rewrite the problem as
\begin{equation}
		\begin{aligned}
			&\text { minimize } \alpha \cdot \mu\\
			&\text { subject to } \|\alpha \cdot \nu\|^2=1
	\end{aligned}
\end{equation}
By setting the Lagrangian function associated with the problem
\begin{equation}
	\mathcal{L}\left(\alpha, \lambda\right)=\alpha \cdot \mu-\lambda\left(\|\alpha \cdot \nu\|^2-1\right) \, ,
\end{equation}
we obtain the first order conditions
\begin{eqsys}
	\frac{\partial \mathcal{L}}{\partial \alpha}=\mu -2\lambda \Sigma \cdot \alpha=0 \\
	\frac{\partial \mathcal{L}}{\partial \lambda}= \|\alpha \cdot \nu\|^2-1=0
	\, ,
\end{eqsys}
Then, by applying simple algebra, we obtain the analytical form of the free optimal strategy
\begin{equation}
	\alpha^* = \pm \frac{\Sigma^{-1} \cdot \mu}{\|(\Sigma^{-1} \cdot \mu)\cdot \nu\|}
\end{equation}
We take the minus sign to get the minimum value of the TVS local drift. 
\end{proof}
\subsubsection{Active Asset (or Bang Bang) Solution}
A closed form solution to the minimization of the local drift correction \eqref{eq:BS_optimal_strategy} can also be derived in the common case that all costs of carry are nonnegative and the only constraint on portfolio weights is nonnegativity, which would mean a long only strategy by the fund manager.

@@

In full generality we suppose that the risky securities follows a generic multidimensional diffusive process $X_t$ given by the following SDE
\begin{equation}
		dX_t = M(X_t)dt + \Sigma(X_t) \cdot dW_t.
\end{equation}
The TVS dynamics is given as follows 
\begin{equation}
	\frac{dI_t}{I_t} = \left(r_t(X_t) -  \frac{\bar{\sigma} \alpha_t}{\|\alpha_t \cdot \nu_t(X_t)\|}  \cdot \mu_t(X_t)    \right)dt +\frac{\bar{\sigma}\alpha_t }{\|\alpha_t \cdot \nu_t(X_t) \|} \cdot \nu_t(X_t) \cdot dW_t
\end{equation}
Given $X := X_t$ and $I := I_t$, we can write the HJB equation for $V := V(t,I,X)$ as follows
\begin{equation}
	\begin{aligned}
		\frac{\partial V}{\partial t} + \max_\alpha  \left\{   \left(r_t(X) - \bar{\sigma}\frac{\alpha \cdot \mu_t(X)}{\|\alpha \cdot \nu_t(X)\|}\right) I \frac{\partial V}{\partial I} + (\nabla_X V) \cdot M(X) + \frac{1}{2} \bar{\sigma}^2I^2 \frac{\partial^2V}{\partial I^2}  \right.\\ 
		\left. +\frac{1}{2}\Tr\left(\tr{\Sigma(X)}(H_XV)\Sigma(X)\right)  + (\nabla_{X,I}V) \cdot \Sigma(X) \cdot \left(I\bar{\sigma}\frac{\alpha \cdot \nu_t(X)}{\|\alpha \cdot \nu_t(X)\|}\right)\right\} = 0,
	\end{aligned}
\end{equation}
where $\Tr(A)$ is the trace operator of $A$, $\nabla_X V$ the gradient of $V$ w.r.t.~$X$, $H_XV$ the Hessian matrix of $V$ w.r.t.~$X$ and $\nabla_{X,I}V$ is the vector defined by:
\begin{equation}
	\nabla_{X,I} V= \tr{\left(\frac{\partial^2 V}{\partial X^1 \partial I},\dots, \frac{\partial^2 V}{\partial X^n \partial I} \right)}.
\end{equation}
We take out from the maximum operator all the elements that do not depend on the risky allocation strategy $\alpha$
\begin{equation}
	\begin{aligned}
		\frac{\partial V}{\partial t} +   r_t(X) I \frac{\partial V}{\partial I} + (\nabla_X V) \cdot M(X) + \frac{1}{2} \bar{\sigma}^2I^2 \frac{\partial^2V}{\partial I^2} + \frac{1}{2}\Tr\left(\tr{\Sigma(X)}(H_XV)\Sigma(X)\right)  \\ 
		+\max_\alpha  \left\{  -\frac{\partial V}{\partial I} I  \bar{\sigma}\frac{\alpha \cdot \mu_t(X)}{\|\alpha \cdot \nu_t(X)\|} + 
		(\nabla_{X,I}V) \cdot \Sigma(X) \cdot \left(I\bar{\sigma}\frac{\alpha \cdot \nu_t(X)}{\|\alpha \cdot \nu_t(X)\|}\right)\right\} = 0.
		\end{aligned}
\label{eq:HJB_general}\end{equation}
\Cref{eq:HJB_general} represent the Hamilton-Jacobi-Bellman equation describing the TVS dynamic problem for a generic dynamics of the risky securities underlying the portfolio.
In our work we have that\footnote{We use $\circ$ for the element-wise product between two vectors and between vector and matrix.} $X_t = S_t$, $M(X_t) \coloneqq (r_t\mathds{1} - \mu_t)\circ S_t$ and $\Sigma(X_t) = S_t \circ \nu_t$.

If we assume a time-dependent BS dynamics for the risky equities ($\mu_t$, $r_t$ and $\nu_t$ deterministic), then $V = V(t,I)$ and all the derivatives w.r.t $X$ are all zero. Thus the reduced HJB is
\begin{equation}
	\frac{\partial V}{\partial t} +  I r(t)\frac{\partial V}{\partial I}  + \frac{1}{2}\bar{\sigma}^2I^2\frac{\partial^2V}{\partial I^2} + \max_\alpha \left\{ -\frac{\partial V}{\partial I} I  \bar{\sigma}\frac{\alpha \cdot \mu(t)}{\|\alpha \cdot \nu(t)\|}\right\} = 0,
\end{equation}
so that if the payoff is increasing in $I_T$ by homogeneity of the SDE we get that $V$ is increasing and so the solution is given by
\begin{equation}
 \alpha^*(t) = \argmin_\alpha \frac{\alpha \cdot \mu(t)}{\|\alpha \cdot \nu(t)\|}
	\end{equation}
which is the same result expressed in \cref{eq:BS_optimal_strategy}. If the payoff is decreasing in $I_T$ then the solution will be the $\argmax$.

@@

\clearpage
\section{Numerical Investigations}\label{sec:Numerical_results}
In this section we present the numerical results of our work. In particular we focus our analysis on an European option on the TVS with at-the-money strike considering two different models for the dynamics followed by the risky securities: Black and Scholes and local volatility. In \Cref{subsec:BS_model} we have proved that under the BS dynamics the solution of the control problem is given by \Cref{eq:BS_optimal_strategy}. We want to take advantage of this \textit{a priori} solution as a benchmark to gather evidence on the robustness of our RL approach and to check if our analytical result is correct. Moreover, the availability of this faster solution allows us to identify the key parameters of the ANN, PPO and the action parameterization, and to analyze how they impact on the final results and performances.  
Then we dedicate a full section to the LV case where the dynamical control problem cannot be solved without the RL agent.

@@

\subsection{Black and Scholes: Hyper-Parameters Fine Tuning}
We use the BS environment as toy model to understand which parameters of the RL algorithm play key roles in the training and testing phase. We also study which action parameterization is most effective in the case of a long only allocation strategy $\alpha$. We consider a target volatility call option with the following contract details between bank and fund manager
\begin{equation}\nonumber
	K = 1 \, \text{EUR}, \quad T=1 \, \text{yr}, \quad \bar{\sigma} = 5\%.
\end{equation}
We test the action parameterizations discussed in \Cref{subsec:policy_action} increasing the number $n$ of equities in the basket; we have evidence that the $(n-1)$-sphere parameterization  \eqref{eq:euler_angles} suffers of labelling problems when $n>2$. In fact if we consider a spherical case ($n=3$), \cref{eq:euler_angles} introduces an asymmetry in the action space (\Cref{fig:n_sphere_parameterization} on the left) that translates into the choice by the agent of sub-optimal policies during the training phase. Because of that we privilege the action parameterization through normalization \eqref{eq:normalized_action}. 
 In \Cref{fig:n_sphere_parameterization} on the right we provide the learning curve of a $5\times8$ ANN\footnote{We use the notation $L\times N$ to describe a FFNN with $L$ hidden layers each made up of $N$ neurons.} with $tanh$ as activation function, $\gamma=1$, $c_1=0.7$ and $c_2=0$. This learning curve, just like all the ones we will show in this work, is the best result in term of reward on the last $10^7$ episodes among four simulations with different initial seed for the weights $\theta$. This is done since the objective function is not convex.

@@

 We adopt as learning strategy the PPO algoritm developed in \textcite{ppo} and \textcite{gae}. We have 
\end{comment}
\cleardoublepage
\printbibliography
\end{document}

>>>>

\usepackage[]{hyperref}
\hypersetup{
	colorlinks=true,breaklinks=true,
	urlcolor= mydarkblue,linkcolor= mydarkblue,citecolor= mydarkgreen,
	pdfauthor={Daluiso, R., Nastasi, E., Pallavicini, A., Polo, S.},
}
\usepackage[backend=biber,style=authoryear-comp,maxcitenames=2,sorting=nyt,natbib=true,sortcites=false]{biblatex}
\setlength\bibitemsep{1.3\itemsep}

\addbibresource{Bib.bib}
%\usepackage{natbib}
%\setcitestyle{authoryear,open={(},close={)}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{remark}{Remark}[section]
\newtheorem{example}[remark]{Example}

@@

\author{
	Roberto Daluiso\thanks{Intesa Sanpaolo Milan, roberto.daluiso@intesasanpaolo.com}
	\and 
	Emanuele Nastasi\thanks{Marketz S.p.A., emanuele.nastasi@marketz.eu}
	\and Andrea Pallavicini\thanks{Imperial College London and Intesa Sanpaolo Milan, a.pallavicini@imperial.ac.uk} 
	\and
	Stefano Polo\thanks{Intesa Sanpaolo Milan, stefano.polo@intesasanpaolo.com} 
}

@@

\begin{abstract}
\noindent In recent years Reinforcement Learning (RL) techniques are gaining popularity in the field of quantitative finance since they are proving to be an efficient way to solve  high-dimensional optimal control problems. 
Our research project is devoted to apply RL to price derivative contracts on target volatility strategies (TVSs), portfolios of risky assets and a risk-free one dynamically rebalanced in order to keep the realized volatility of the portfolio on a certain level. The uncertainty in the TVS risky portfolio composition along with the difference in hedging costs for each component requires to solve a stochastic control problem to evaluate the option prices. The topic of hedging costs is a novelty never dealt in the TVS literature and we provide a formal description of the entire control problem. We tackle the problem by implementing a RL algorithm to determine the optimal risky portfolio composition leading to the most conservative option price.  We investigate the problem for two models of the risky asset dynamics: time-dependent Black and Scholes and local volatility. In the first case we prove the existence of an analytical solution of the problem; a result that we use as benchmark to perform a series of fine-tuning of the RL hyper-parameters. At the end we provide numerical results for the local volatility model, for which an \textit{a priori} solution is not available.
\end{abstract}

\smallskip
\noindent
\textbf{JEL classification codes:} C63, C45, C61, G11, G13.\\
\textbf{AMS classification codes:} 65C05, 91G20, 91G60.\\
\textbf{Keywords:} Reinforcement Learning, Hedging Costs, Funding Costs, Proximal Policy Optimization, Target Volatility, Asset Allocation. 
\newpage
\tableofcontents
 \mbox{}
 \vfill
 \noindent \small The opinions here expressed are solely those of the authors and do not represent in any way those of their employers.

\newpage
\pagestyle{myheadings} \markboth{}{{\footnotesize Daluiso, Pallavicini, Nastasi, Polo, Reinforcement Learning for Options on Target Volatility Funds}}
\section{Introduction}
In the recent years portfolio managers are exposed to very low interest rates and quickly changing market volatilities. An effective solution to control risks under such environment is given by target volatility strategies (TVSs) (also known as constant volatility targeting) which are able to preserve the portfolio at a predetermined level of volatility. A TVS is  a  portfolio  of  risky  assets  (typically  equities)  and  a  risk-free  asset dynamically re-balanced with the aim of maintaining the overall portfolio volatility level closed to some target value. This products were initially offered in the Asian markets, see for instance the reports of \textcite{Chew} and \textcite{XUE} which highlight the pros and cons for investors, to be adopted in the following years in many other markets in North America and Europe as depicted in \textcite{Morrison}.

@@

In this contribution we face a new topic never dealt with in the TVS literature: the funding costs coming from hedging the risky assets underlying the TVS. We consider the point of view of a bank selling a call option to a portfolio manager as protection on the capital invested in a TVS. The portfolio manager has the freedom of changing the relative weights of the risky assets during the life of the TVS. Since the risky assets have different hedging costs, the bank shall adjust the price of the protection to include them in the worst-case scenario, i.e. the most expensive strategy from the financing costs point of view. Hence, the pricing problem becomes a dynamical control problem over the risky portfolio composition. In our contribution we describe the dynamical control problem, and we derive an analytical solution in the simple case of the risky assets driven by a Black-Scholes (BS) model \citep{BlackScholes}. Then, we tackle the problem in the general case of a local volatility (LV) model \citep{Dupire94pricingwith,LVfirst} by using Reinforcement Learning (RL) algorithms, in particular the proximal policy optimization developed in \textcite{ppo}.
\\

The paper is organized ad follows. In \Cref{sec:TVS} we describe the dynamics of a TVS in presence of valuation adjustments. Then, in \Cref{sec:Derivative} we study the case of derivative contracts on TVSs, we describe the dynamical control problem giving the expression of the Hamilton-Jacobi-Bellman equation and we derive the analytical results for plain vanilla options under BS dynamics. In \Cref{sec:RL} we illustrate how we have applied RL to solve the dynamic control problem, giving a description of the algorithm we have built. We conclude the paper with \Cref{sec:Numerical_results} where we present the numerical results obtained in this work for the Black and Scholes model and the local volatility one.
\\

A part of this work has been developed during the Master thesis of one of us (SP), where we had the opportunity to collaborate with Marco Bianchetti and Diego Pierluigi Giovannini from Intesa Sanpaolo, Milan, that we tanks. Moreover we wish to thank the Italian computing centre Cineca, which approved our ISCRA C project and provided us the high-computing resources of Marconi100 for the numerical simulations of this contribution.

@@

where $\Phi$ is the payoff function of the derivative. We start by introducing the Markovian projection of the dynamics followed by $I_t$. We name it $I_t^{\text{MP}}$, and we get by applying the Gy\"ongy Lemma \citep{Gyongy}

@@

In a more general settings we are not able to find an explicit solution. A proof of the existence of the solution in a general setting is missing. This is a stochastic optimal control problem where by homogeneity we can suppose that $\alpha_t$ lives in a compact domain, namely (a subset of) the unit simplex. A least if $r_t$, $\mu^i_t$ and $\nu^i_t$ are (uniformly) bounded, and if the eigenvalues of $\nu_t$ are (uniformly) bounded away from zero, then drift and diffusion should be uniformly Lipschitz in $\alpha_t$, and classical theorems should exist \citep{StandardTheorems}. 

@@

\begin{proposition}
When the underlying securities follow a Black and Scholes (BS) model with deterministic rates, the optimal strategy for a non-decreasing European payoff consists in minimizing the local drift function, independently of the current state $I_t$
\begin{equation}
    \alpha^*(t) \coloneqq \argmin_\alpha \frac{\alpha \cdot \mu(t)}{\|\alpha \cdot \nu(t) \|}.
\label{eq:BS_optimal_strategy}
\end{equation}
Analogously, the optimal strategy for a non-increasing European payoff consists in maximizing the local drift function:
\begin{equation}
    \alpha^*(t) \coloneqq \argmax_\alpha \frac{\alpha \cdot \mu(t)}{\|\alpha \cdot \nu(t) \|}.
\label{eq:BS_optimal_strategy_put}
\end{equation}
\end{proposition}
The absence of stochastic elements in \Cref{eq:BS_optimal_strategy} makes the optimal strategy known \textit{a priori} with no simulation needed; in fact one can solve the optimization problem once for all $\forall t \in \mathcal{T}$ just looking at the market data $\mu(t)$ and $\nu(t)$ for the securities. Once $\alpha^*$ is known, then one can price the payoff by the following BS formula
\begin{equation}
	V_0^{\text{BS}} = BS(F^{\text{TVS}}(0,T;\alpha^*), K, T, \bar{\sigma}, D(0,T;\zeta)),
\label{eq:bs_closed_solution}\end{equation}
where $F^{\text{TVS}}(t,T;\alpha)$ is the TVS forward curve defined by
\begin{equation}
	F^{\text{TVS}}(t,T;\alpha^*) = I_t \exp\left[\int_t^T\left(r(u)-\ell_{\alpha^*} \left(u\right)\right)du\right] \quad \text{with} \quad \ell_{\alpha^*}(t)=\bar{\sigma}   \frac{\mu\left(t\right) \cdot \alpha^*\left(t\right)}{\|\alpha^*\left(t\right) \cdot \nu\left(t\right)\|},
\end{equation}
while $BS(F,K,T,\sigma,D)$ is the standard BS formula for an European option with forward curve $F$, strike $K$, time to maturity $T$, volatility $\sigma$ and discount factor $D$.  
\subsubsection{Optimal Strategy Closed Form}
In absence of constraints on the allocation strategy, we are able to derive a closed form solution to the BS problem \eqref{eq:BS_optimal_strategy}
\begin{lemma}
	Let be $\mu, \alpha \in \mathbb{R}^n$, $\nu \in \mathbb{R}^{n\times n}$ be a full rank matrix and $\Sigma\coloneqq \nu \tr{\nu}$. Then the closed solution of the optimization problem \eqref{eq:BS_optimal_strategy} is 
	\begin{equation}
			\alpha^* = - \frac{\Sigma^{-1} \cdot \mu}{\|(\Sigma^{-1} \cdot \mu)\cdot \nu\|}
	\end{equation}
\end{lemma}
\begin{proof}
Since the argument of the minimum \eqref{eq:BS_optimal_strategy} is zero-homogeneous, then we can rewrite the problem as
\begin{equation}
		\begin{aligned}
			&\text { minimize } \alpha \cdot \mu\\
			&\text { subject to } \|\alpha \cdot \nu\|^2=1
	\end{aligned}
\end{equation}
By setting the Lagrangian function associated with the problem
\begin{equation}
	\mathcal{L}\left(\alpha, \lambda\right)=\alpha \cdot \mu-\lambda\left(\|\alpha \cdot \nu\|^2-1\right) \, ,
\end{equation}
we obtain the first order conditions
\begin{eqsys}
	\frac{\partial \mathcal{L}}{\partial \alpha}=\mu -2\lambda \Sigma \cdot \alpha=0 \\
	\frac{\partial \mathcal{L}}{\partial \lambda}= \|\alpha \cdot \nu\|^2-1=0
	\, ,
\end{eqsys}
Then, by applying simple algebra, we obtain the analytical form of the free optimal strategy
\begin{equation}
	\alpha^* = \pm \frac{\Sigma^{-1} \cdot \mu}{\|(\Sigma^{-1} \cdot \mu)\cdot \nu\|}
\end{equation}
We take the minus sign to get the minimum value of the TVS local drift. 
\end{proof}
\subsubsection{Active Asset (or Bang Bang) Solution}
A closed form solution to the minimization of the local drift correction \eqref{eq:BS_optimal_strategy} can also be derived in the common case that all costs of carry are nonnegative and the only constraint on portfolio weights is nonnegativity, which would mean a long only strategy by the fund manager.

@@

In full generality we suppose that the risky securities follows a generic multidimensional diffusive process $X_t$ given by the following SDE
\begin{equation}
		dX_t = M(X_t)dt + \Sigma(X_t) \cdot dW_t.
\end{equation}
The TVS dynamics is given as follows 
\begin{equation}
	\frac{dI_t}{I_t} = \left(r_t(X_t) -  \frac{\bar{\sigma} \alpha_t}{\|\alpha_t \cdot \nu_t(X_t)\|}  \cdot \mu_t(X_t)    \right)dt +\frac{\bar{\sigma}\alpha_t }{\|\alpha_t \cdot \nu_t(X_t) \|} \cdot \nu_t(X_t) \cdot dW_t
\end{equation}
Given $X := X_t$ and $I := I_t$, we can write the HJB equation for $V := V(t,I,X)$ as follows
\begin{equation}
	\begin{aligned}
		\frac{\partial V}{\partial t} + \max_\alpha  \left\{   \left(r_t(X) - \bar{\sigma}\frac{\alpha \cdot \mu_t(X)}{\|\alpha \cdot \nu_t(X)\|}\right) I \frac{\partial V}{\partial I} + (\nabla_X V) \cdot M(X) + \frac{1}{2} \bar{\sigma}^2I^2 \frac{\partial^2V}{\partial I^2}  \right.\\ 
		\left. +\frac{1}{2}\Tr\left(\tr{\Sigma(X)}(H_XV)\Sigma(X)\right)  + (\nabla_{X,I}V) \cdot \Sigma(X) \cdot \left(I\bar{\sigma}\frac{\alpha \cdot \nu_t(X)}{\|\alpha \cdot \nu_t(X)\|}\right)\right\} = 0,
	\end{aligned}
\end{equation}
where $\Tr(A)$ is the trace operator of $A$, $\nabla_X V$ the gradient of $V$ w.r.t.~$X$, $H_XV$ the Hessian matrix of $V$ w.r.t.~$X$ and $\nabla_{X,I}V$ is the vector defined by:
\begin{equation}
	\nabla_{X,I} V= \tr{\left(\frac{\partial^2 V}{\partial X^1 \partial I},\dots, \frac{\partial^2 V}{\partial X^n \partial I} \right)}.
\end{equation}
We take out from the maximum operator all the elements that do not depend on the risky allocation strategy $\alpha$
\begin{equation}
	\begin{aligned}
		\frac{\partial V}{\partial t} +   r_t(X) I \frac{\partial V}{\partial I} + (\nabla_X V) \cdot M(X) + \frac{1}{2} \bar{\sigma}^2I^2 \frac{\partial^2V}{\partial I^2} + \frac{1}{2}\Tr\left(\tr{\Sigma(X)}(H_XV)\Sigma(X)\right)  \\ 
		+\bar{\sigma}I\max_\alpha  \left\{  -\frac{\partial V}{\partial I}   \frac{\alpha \cdot \mu_t(X)}{\|\alpha \cdot \nu_t(X)\|} + 
		(\nabla_{X,I}V) \cdot \Sigma(X) \cdot \left(\frac{\alpha \cdot \nu_t(X)}{\|\alpha \cdot \nu_t(X)\|}\right)\right\} = 0.
		\end{aligned}
\label{eq:HJB_general}\end{equation}
\Cref{eq:HJB_general} represent the Hamilton-Jacobi-Bellman equation describing the TVS dynamic problem for a generic dynamics of the risky securities underlying the portfolio.
In our work we have that\footnote{We use $\circ$ for the element-wise product between two vectors and between vector and matrix.} $X_t = S_t$, $M(X_t) \coloneqq (r_t\mathds{1} - \mu_t)\circ S_t$ and $\Sigma(X_t) = S_t \circ \nu_t$.

If we assume a time-dependent BS dynamics for the risky equities ($\mu_t$, $r_t$ and $\nu_t$ deterministic), then $V = V(t,I)$ and all the derivatives w.r.t $X$ are all zero. Thus the reduced HJB is
\begin{equation}
	\frac{\partial V}{\partial t} +  I r(t)\frac{\partial V}{\partial I}  + \frac{1}{2}\bar{\sigma}^2I^2\frac{\partial^2V}{\partial I^2} + \bar{\sigma}I\max_\alpha \left\{ -\frac{\partial V}{\partial I}   \frac{\alpha \cdot \mu(t)}{\|\alpha \cdot \nu(t)\|}\right\} = 0,
\end{equation}
so that if the payoff is increasing in $I_T$ by homogeneity of the SDE we get that $V$ is increasing and so the solution is given by
\begin{equation}
 \alpha^*(t) = \argmin_\alpha \frac{\alpha \cdot \mu(t)}{\|\alpha \cdot \nu(t)\|}
	\end{equation}
which is the same result expressed in \cref{eq:BS_optimal_strategy}. If the payoff is decreasing in $I_T$ then the solution will be the $\argmax$.

@@

\section{Numerical Investigations}\label{sec:Numerical_results}
In this section we present the numerical results of our work. In particular we focus our analysis on an European option on the TVS with at-the-money strike considering two different models for the dynamics followed by the risky securities: Black and Scholes and local volatility. In \Cref{subsec:BS_model} we have proved that under the BS dynamics the solution of the control problem is given by \Cref{eq:BS_optimal_strategy}. We want to take advantage of this \textit{a priori} solution as a benchmark to gather evidence on the robustness of our RL approach and to check if our analytical result is correct. Moreover, the availability of this faster solution allows us to identify the key parameters of the ANN, PPO and the action parameterization, and to analyze how they impact on the final results and performances.  
Then we dedicate a full section to the LV case where the dynamical control problem cannot be solved without the RL agent.

@@

\subsection{Black and Scholes: Hyper-Parameters Fine Tuning}
We use the BS environment as toy model to understand which parameters of the RL algorithm play key roles in the training and testing phase. We also study which action parameterization is most effective in the case of a long only allocation strategy $\alpha$. We consider a target volatility call option with the following contract details between bank and fund manager
\begin{equation}\nonumber
	I_0 = K = 1 \, \text{EUR}, \quad T=1 \, \text{yr}, \quad \bar{\sigma} = 5\%.
\end{equation}
We test the action parameterizations discussed in \Cref{subsec:policy_action} increasing the number $n$ of equities in the basket; we have evidence that the $(n-1)$-sphere parameterization  \eqref{eq:euler_angles} suffers of labelling problems when $n>2$. In fact if we consider a spherical case ($n=3$), \cref{eq:euler_angles} introduces an asymmetry in the action space (\Cref{fig:n_sphere_parameterization} on the left) that translates into the choice by the agent of sub-optimal policies during the training phase. Because of that we privilege the action parameterization through normalization \eqref{eq:normalized_action}. 
 In \Cref{fig:n_sphere_parameterization} on the right we provide the learning curve of a $5\times8$ ANN\footnote{We use the notation $L\times N$ to describe a FFNN with $L$ hidden layers each made up of $N$ neurons.} with $\tanh$ as activation function, $\gamma=1$, $c_1=0.7$ and $c_2=0$. This learning curve, just like all the ones we will show in this work, is the best result in term of reward on the last $10^7$ episodes among four simulations with different initial seed for the weights $\theta$. This is done since the objective function is not convex.

@@

 We adopt as learning strategy the PPO algoritm developed in \textcite{ppo} and \textcite{gae}. We have 
\end{comment}
\\

\printbibliography
\end{document}
